---
title: "precisionFDA"
author: "Pallavi Misra"
date: "6/18/2020"
output:
  html_document: default
  pdf_document: default
---

```{r library, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### __Dependencies and Requirements__
* R version 4.0.0
* RTools40
* PostgreSQL12
* Download all CDM version 5 vocabularies from Athena.ohdsi.org

#### Packages to install

```{r setup, eval=TRUE}
install.packages('devtools')
install.packages('bit64')
install.packages("DatabaseConnector")
install.packages("usethis")
devtools::install_github("OHDSI/ETL-Synthea")
install.packages("drat")
drat::addRepo("OHDSI")
install.packages("PatientLevelPrediction")
install.packages("FeatureExtraction") 
install.packages("Andromeda") 
install.packages("dplyr") 
install.packages("Cyclops")
```

#### Load Libraries

```{r libraries, eval=TRUE}
library(usethis)
library(DatabaseConnector)
library(ETLSyntheaBuilder)
library(SqlRender)
library(devtools) 
library(dplyr)
library(Andromeda)
library(FeatureExtraction) 
library(Cyclops)
library(PatientLevelPrediction)
```

#### Make connection with PostgreSQL

Enter your server details below.

```{r connection details, eval=TRUE}
cd <- DatabaseConnector::createConnectionDetails(
  dbms     = "postgresql", 
  server   = "localhost/postgres", 
  user     = "postgres", 
  password = "****", 
  port     = 5432)
conn <- connect(cd)
```



### Create the Common Data Model for Train Data

#### Create schemas for Common Data Model, vocabularies and csv files
_execute this command only if schemas do not exist, otherwise it will throw an error_

```{r create schemas, eval=TRUE}
sql <- "create schema cdm_precision;
create schema precision_native;"
renderTranslateExecuteSql(conn, sql)
```

#### Drop all existing tables 

```{r drop existing tables, eval=TRUE}
ETLSyntheaBuilder::DropVocabTables(connectionDetails = cd, vocabDatabaseSchema = "cdm_precision")
ETLSyntheaBuilder::DropEventTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision")
ETLSyntheaBuilder::DropSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native")
ETLSyntheaBuilder::DropMapAndRollupTables (connectionDetails = cd, cdmDatabaseSchema = "cdm_precision")
```

#### Import csv files

syntheaFileLoc: location of input csv files

_Delete the Organizations column from encounters.csv_

```{r import csv files, eval=TRUE}
ETLSyntheaBuilder::CreateSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native")
ETLSyntheaBuilder::LoadSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native", syntheaFileLoc = syntheaFileLoc) 
```

#### Create vocabulary tables

```{r create vocabulary tables, eval=TRUE}
ETLSyntheaBuilder::CreateVocabTables(connectionDetails = cd, vocabDatabaseSchema = "cdm_precision")
```

#### Change type of concept_name and concept_synonym_name to text

```{r change type, eval=TRUE}
sql <- "alter table @cdm.concept alter column concept_name type text;
alter table @cdm.concept_synonym alter column concept_synonym_name type text;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")
```

#### Load vocabulary tables

vocabFileLoc: location of vocabulary files

```{r load vocabulary tables, eval=TRUE}
ETLSyntheaBuilder::LoadVocabFromCsv(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision", vocabFileLoc = vocabFileLoc) 
```

#### Build the Common Data Model

```{r CDM Builder, eval=TRUE}
ETLSyntheaBuilder::CreateVocabMapTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision")
ETLSyntheaBuilder::CreateEventTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision")
ETLSyntheaBuilder::CreateVisitRollupTables(cd, cdmDatabaseSchema = "cdm_precision", syntheaDatabaseSchema = "precision_native")
ETLSyntheaBuilder::LoadEventTables(cd, cdmDatabaseSchema = "cdm_precision", syntheaDatabaseSchema = "precision_native", vocabDatabaseSchema = "cdm_precision")
```

#### Change names of a few columns

```{r change names, eval=TRUE}
sql <- "alter table cdm_precision.condition_era rename column condition_era_start_datetime to condition_era_start_date;
alter table cdm_precision.condition_era rename column condition_era_end_datetime to condition_era_end_date;
alter table cdm_precision.drug_era rename column drug_era_start_datetime to drug_era_start_date;
alter table cdm_precision.drug_era rename column drug_era_end_datetime to drug_era_end_date;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")
```



### Create the Common Data Model for Test Data

#### Create schemas for Common Data Model, vocabularies and csv files
_execute this command only if schemas do not exist, otherwise it will throw an error_

```{r create schemas, eval=TRUE}
sql <- "create schema cdm_precision_test;
create schema precision_native_test;"
renderTranslateExecuteSql(conn, sql)
```

#### Drop all existing tables 

```{r drop existing tables, eval=TRUE}
ETLSyntheaBuilder::DropVocabTables(connectionDetails = cd, vocabDatabaseSchema = "cdm_precision_test")
ETLSyntheaBuilder::DropEventTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision_test")
ETLSyntheaBuilder::DropSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native_test")
ETLSyntheaBuilder::DropMapAndRollupTables (connectionDetails = cd, cdmDatabaseSchema = "cdm_precision_test")
```

#### Import csv files

syntheaFileLoc: location of input csv files

_Delete the Organizations column from encounters.csv_

```{r import csv files, eval=TRUE}
ETLSyntheaBuilder::CreateSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native_test")
ETLSyntheaBuilder::LoadSyntheaTables(connectionDetails = cd, syntheaDatabaseSchema = "precision_native_test", syntheaFileLoc = syntheaFileLoc) 
```

#### Create vocabulary tables

```{r create vocabulary tables, eval=FALSE}
ETLSyntheaBuilder::CreateVocabTables(connectionDetails = cd, vocabDatabaseSchema = "cdm_precision_test")
```

#### Change type of concept_name and concept_synonym_name to text

```{r change type, eval=TRUE}
sql <- "alter table @cdm.concept alter column concept_name type text;
alter table @cdm.concept_synonym alter column concept_synonym_name type text;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")
```

#### Load vocabulary tables

vocabFileLoc: location of vocabulary files

```{r load vocabulary tables, eval=FALSE}
ETLSyntheaBuilder::LoadVocabFromCsv(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision_test", vocabFileLoc = vocabFileLoc) 
```

#### Build the Common Data Model

```{r CDM Builder, eval=TRUE}
ETLSyntheaBuilder::CreateVocabMapTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision_test")
ETLSyntheaBuilder::CreateEventTables(connectionDetails = cd, cdmDatabaseSchema = "cdm_precision_test")
ETLSyntheaBuilder::CreateVisitRollupTables(cd, cdmDatabaseSchema = "cdm_precision_test", syntheaDatabaseSchema = "precision_native_test")
ETLSyntheaBuilder::LoadEventTables(cd, cdmDatabaseSchema = "cdm_precision_test", syntheaDatabaseSchema = "precision_native_test", vocabDatabaseSchema = "cdm_precision_test")
```

#### Change names of a few columns

```{r change names, eval=TRUE}
sql <- "alter table cdm_precision_test.condition_era rename column condition_era_start_datetime to condition_era_start_date;
alter table cdm_precision_test.condition_era rename column condition_era_end_datetime to condition_era_end_date;
alter table cdm_precision_test.drug_era rename column drug_era_start_datetime to drug_era_start_date;
alter table cdm_precision_test.drug_era rename column drug_era_end_datetime to drug_era_end_date;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")
```



### Train data- Make Cohorts, set Covariate, extract plpData, make Population and Run Model

#### Make cohort table in cdm_precision schema

```{r create cohort table, eval=TRUE}
sql <- "drop table if exists @cdm.cohort;
CREATE TABLE @cdm.cohort (  subject_id bigint,  cohort_start_date date, cohort_end_date date,  cohort_definition_id integer);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")
```

#### Cohort 1: All patients

```{r create cohort 1, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (1 AS INT) AS cohort_definition_id
FROM #diagnoses;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)
```

#### Cohort 2: COVID-19 patients

```{r create cohort 2, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
     WHERE ancestor_concept_id = 37311061 -- COVID-19
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (2 AS INT) AS cohort_definition_id
FROM #diagnoses;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)
```

#### Cohort 3: Deceased patients

```{r create cohort 3, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
     WHERE ancestor_concept_id = 37311061 -- COVID-19
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "SELECT person_id AS subject_id,
 death_datetime
INTO #diagnoses_one
FROM @cdm.person
WHERE death_datetime is not NULL;"

renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "select * 
into #diagnoses_two
from (Select abc.subject_id, abc.death_datetime as cohort_start_date, date('2070-01-01') as cohort_end_date
From #diagnoses_one abc
Inner Join #diagnoses on abc.subject_id=#diagnoses.subject_id) as whatever;"

renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (3 AS INT) AS cohort_definition_id
FROM #diagnoses_two;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)

sql <- "TRUNCATE TABLE #diagnoses_one;
DROP TABLE #diagnoses_one;"
renderTranslateExecuteSql(conn, sql)

sql <- "TRUNCATE TABLE #diagnoses_two;
DROP TABLE #diagnoses_two;"
renderTranslateExecuteSql(conn, sql)
```

#### Cohort 4: Ventilation patients

```{r create cohort 4, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
     WHERE ancestor_concept_id = 37311061 -- COVID-19
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "SELECT person_id AS subject_id
INTO #diagnoses_one
FROM @cdm.procedure_occurrence
WHERE procedure_source_value = '26763009' -- Ventilation
;"

renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "select abc.subject_id, abc.cohort_start_date, abc.cohort_end_date 
into #diagnoses_two
From #diagnoses abc
Inner Join #diagnoses_one on abc.subject_id=#diagnoses_one.subject_id;"

renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (4 AS INT) AS cohort_definition_id
FROM #diagnoses_two;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)

sql <- "TRUNCATE TABLE #diagnoses_one;
DROP TABLE #diagnoses_one;"
renderTranslateExecuteSql(conn, sql)

sql <- "TRUNCATE TABLE #diagnoses_two;
DROP TABLE #diagnoses_two;"
renderTranslateExecuteSql(conn, sql)
```

#### Covariate settings

```{r covariate settings COVID, eval=TRUE}
covSettings_COVID <- createCovariateSettings(useDemographicsGender = TRUE,
                                    useDemographicsAge = TRUE,
                                    useDemographicsRace = TRUE,
                                    useDemographicsEthnicity = TRUE,
                                    useDemographicsIndexYear = TRUE,
                                    useConditionOccurrenceAnyTimePrior = TRUE,
                                    useConditionOccurrencePrimaryInpatientAnyTimePrior = TRUE,
                                    useConditionEraAnyTimePrior = TRUE,
                                    useConditionGroupEraAnyTimePrior = TRUE,
                                    useDrugExposureAnyTimePrior = TRUE,
                                    useDrugEraAnyTimePrior = TRUE,
                                    useDrugGroupEraAnyTimePrior = TRUE,
                                    useProcedureOccurrenceAnyTimePrior = TRUE,
                                    useDeviceExposureAnyTimePrior = TRUE,
                                    useMeasurementAnyTimePrior = TRUE,
                                    useMeasurementValueAnyTimePrior = TRUE,
                                    useMeasurementRangeGroupAnyTimePrior = TRUE,
                                    useObservationAnyTimePrior = TRUE,
                                    useCharlsonIndex = TRUE,
                                    useDcsi = TRUE,
                                    useChads2 = TRUE,
                                    useChads2Vasc = TRUE,
                                    useHfrs = TRUE,
                                    useDistinctConditionCountLongTerm = TRUE,
                                    useDistinctIngredientCountLongTerm = TRUE,
                                    useDistinctProcedureCountLongTerm = TRUE,
                                    useDistinctMeasurementCountLongTerm = TRUE,
                                    useDistinctObservationCountLongTerm = TRUE,
                                    useVisitCountLongTerm = TRUE,
                                    useVisitConceptCountLongTerm = TRUE,
                                    longTermStartDays = -365,
                                    excludedCovariateConceptIds = ####################,
                                    endDays = 0)
```

```{r covariate settings Death, eval=TRUE}
covSettings_COVIDDeath <- createCovariateSettings(useDemographicsGender = TRUE,
                                    useDemographicsAge = TRUE,
                                    useDemographicsRace = TRUE,
                                    useDemographicsEthnicity = TRUE,
                                    useDemographicsIndexYear = TRUE,
                                    useConditionOccurrenceAnyTimePrior = TRUE,
                                    useConditionOccurrencePrimaryInpatientAnyTimePrior = TRUE,
                                    useConditionEraAnyTimePrior = TRUE,
                                    useConditionGroupEraAnyTimePrior = TRUE,
                                    useDrugExposureAnyTimePrior = TRUE,
                                    useDrugEraAnyTimePrior = TRUE,
                                    useDrugGroupEraAnyTimePrior = TRUE,
                                    useProcedureOccurrenceAnyTimePrior = TRUE,
                                    useDeviceExposureAnyTimePrior = TRUE,
                                    useMeasurementAnyTimePrior = TRUE,
                                    useMeasurementValueAnyTimePrior = TRUE,
                                    useMeasurementRangeGroupAnyTimePrior = TRUE,
                                    useObservationAnyTimePrior = TRUE,
                                    useCharlsonIndex = TRUE,
                                    useDcsi = TRUE,
                                    useChads2 = TRUE,
                                    useChads2Vasc = TRUE,
                                    useHfrs = TRUE,
                                    useDistinctConditionCountLongTerm = TRUE,
                                    useDistinctIngredientCountLongTerm = TRUE,
                                    useDistinctProcedureCountLongTerm = TRUE,
                                    useDistinctMeasurementCountLongTerm = TRUE,
                                    useDistinctObservationCountLongTerm = TRUE,
                                    useVisitCountLongTerm = TRUE,
                                    useVisitConceptCountLongTerm = TRUE,
                                    longTermStartDays = -365,
                                    endDays = 0)
```

```{r covariate settings Vent, eval=TRUE}
covSettings_COVIDVent <- createCovariateSettings(useDemographicsGender = TRUE,
                                    useDemographicsAge = TRUE,
                                    useDemographicsRace = TRUE,
                                    useDemographicsEthnicity = TRUE,
                                    useDemographicsIndexYear = TRUE,
                                    useConditionOccurrenceAnyTimePrior = TRUE,
                                    useConditionOccurrencePrimaryInpatientAnyTimePrior = TRUE,
                                    useConditionEraAnyTimePrior = TRUE,
                                    useConditionGroupEraAnyTimePrior = TRUE,
                                    useDrugExposureAnyTimePrior = TRUE,
                                    useDrugEraAnyTimePrior = TRUE,
                                    useDrugGroupEraAnyTimePrior = TRUE,
                                    useProcedureOccurrenceAnyTimePrior = TRUE,
                                    useDeviceExposureAnyTimePrior = TRUE,
                                    useMeasurementAnyTimePrior = TRUE,
                                    useMeasurementValueAnyTimePrior = TRUE,
                                    useMeasurementRangeGroupAnyTimePrior = TRUE,
                                    useObservationAnyTimePrior = TRUE,
                                    useCharlsonIndex = TRUE,
                                    useDcsi = TRUE,
                                    useChads2 = TRUE,
                                    useChads2Vasc = TRUE,
                                    useHfrs = TRUE,
                                    useDistinctConditionCountLongTerm = TRUE,
                                    useDistinctIngredientCountLongTerm = TRUE,
                                    useDistinctProcedureCountLongTerm = TRUE,
                                    useDistinctMeasurementCountLongTerm = TRUE,
                                    useDistinctObservationCountLongTerm = TRUE,
                                    useVisitCountLongTerm = TRUE,
                                    useVisitConceptCountLongTerm = TRUE,
                                    longTermStartDays = -365,
                                    excludedCovariateConceptIds = ###################3,
                                    endDays = 0)
```

#### Extract cohort data from cdm_precision schema

```{r extract cohort data COVID, eval=TRUE}
plpData_COVID <- getPlpData(connectionDetails = cd,
                      cdmDatabaseSchema = "cdm_precision",
                      cohortDatabaseSchema = "cdm_precision",
                      cohortTable = "cohort",
                      cohortId = 1,
                      covariateSettings = covSettings_COVID,
                      outcomeDatabaseSchema = "cdm_precision",
                      outcomeTable = "cohort",
                      outcomeIds = 2)

```

```{r extract cohort data Death, eval=TRUE}
plpData_COVIDDeath <- getPlpData(connectionDetails = cd,
                      cdmDatabaseSchema = "cdm_precision",
                      cohortDatabaseSchema = "cdm_precision",
                      cohortTable = "cohort",
                      cohortId = 1,
                      covariateSettings = covSettings_COVIDDeath,
                      outcomeDatabaseSchema = "cdm_precision",
                      outcomeTable = "cohort",
                      outcomeIds = 3)

```

```{r extract cohort data Ventilation, eval=TRUE}
plpData_COVIDVent <- getPlpData(connectionDetails = cd,
                      cdmDatabaseSchema = "cdm_precision",
                      cohortDatabaseSchema = "cdm_precision",
                      cohortTable = "cohort",
                      cohortId = 1,
                      covariateSettings = covSettings_COVIDVent,
                      outcomeDatabaseSchema = "cdm_precision",
                      outcomeTable = "cohort",
                      outcomeIds = 4)

```

#### Create a study population

```{r create study population COVID, eval=TRUE}
population_COVID <- createStudyPopulation(plpData = plpData_COVID,
                                    outcomeId = 2,
                                    includeAllOutcomes = TRUE,
                                    firstExposureOnly = FALSE,
                                    washoutPeriod = 0,
                                    removeSubjectsWithPriorOutcome = FALSE,
                                    priorOutcomeLookback = 9999,
                                    requireTimeAtRisk = TRUE,
                                    minTimeAtRisk = 0,
                                    riskWindowStart = 0,
                                    riskWindowEnd = 365)
```

```{r create study population Death, eval=TRUE}
population_COVIDDeath <- createStudyPopulation(plpData = plpData_COVIDDeath,
                                    outcomeId = 3,
                                    includeAllOutcomes = TRUE,
                                    firstExposureOnly = FALSE,
                                    washoutPeriod = 0,
                                    removeSubjectsWithPriorOutcome = FALSE,
                                    priorOutcomeLookback = 9999,
                                    requireTimeAtRisk = TRUE,
                                    minTimeAtRisk = 0,
                                    riskWindowStart = 0,
                                    riskWindowEnd = 365)
```

```{r create study population Ventilation, eval=TRUE}
population_COVIDVent <- createStudyPopulation(plpData = plpData_COVIDVent,
                                    outcomeId = 4,
                                    includeAllOutcomes = TRUE,
                                    firstExposureOnly = FALSE,
                                    washoutPeriod = 0,
                                    removeSubjectsWithPriorOutcome = FALSE,
                                    priorOutcomeLookback = 9999,
                                    requireTimeAtRisk = TRUE,
                                    minTimeAtRisk = 0,
                                    riskWindowStart = 0,
                                    riskWindowEnd = 365)
```

#### Random Forest models settings

```{r Random Forest Model settings, eval=TRUE}
RFModel <- setRandomForest(  mtries = 20,
                             ntrees = 500,
                             maxDepth = 7,
                             varImp = T,
                             seed = 0)
```

#### Random Forest Model- COVID

```{r Random Forest Model COVID, eval=TRUE}
RFResults <- runPlp(population = population_COVID, 
                     plpData = plpData_COVID, 
                     modelSettings = RFModel, 
                     saveDirectory = "C:/Users/misrap/Desktop/COVID/precisionFDA/ML_models",
                     testSplit = 'person',
                     testFraction = 0.25, 
                     nfold = 3, 
                     splitSeed = 0)
```

#### Random Forest Model- COVIDDeath

```{r Random Forest Model Death, eval=TRUE}
RFResults <- runPlp(population = population_COVIDDeath, 
                     plpData = plpData_COVID, 
                     modelSettings = RFModel, 
                     saveDirectory = "C:/Users/misrap/Desktop/COVID/precisionFDA/ML_models",
                     testSplit = 'person',
                     testFraction = 0.25, 
                     nfold = 3, 
                     splitSeed = 0)
```

#### Random Forest Model- COVIDVent

```{r Random Forest Model Ventilation, eval=TRUE}
RFResults <- runPlp(population = population_COVIDVent, 
                     plpData = plpData_COVIDVent, 
                     modelSettings = RFModel, 
                     saveDirectory = "C:/Users/misrap/Desktop/COVID/precisionFDA/ML_models",
                     testSplit = 'person',
                     testFraction = 0.25, 
                     nfold = 3, 
                     splitSeed = 0)
```



### Test data- Make Cohorts, set Covariate, extract plpData, make Population and get Predictions

#### Make Cohorts

#### Make cohort table in cdm_precision_test schema

```{r create cohort table, eval=TRUE}
sql <- "drop table if exists @cdm.cohort;
CREATE TABLE @cdm.cohort (  subject_id bigint,  cohort_start_date date, cohort_end_date date,  cohort_definition_id integer);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")
```

#### Cohort 1: All patients

```{r create cohort 1, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (1 AS INT) AS cohort_definition_id
FROM #diagnoses;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)
```

#### Cohort 2: All patients

```{r create cohort 2, eval=TRUE}
sql <- "SELECT person_id AS subject_id,
 condition_start_date AS cohort_start_date,
 condition_end_date AS cohort_end_date
INTO #diagnoses
FROM @cdm.condition_occurrence
WHERE condition_concept_id IN (
   SELECT descendant_concept_id
   FROM @cdm.concept_ancestor
);"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")

sql <- "INSERT INTO @cdm.cohort (
  subject_id, 
  cohort_start_date,
  cohort_end_date,
  cohort_definition_id
  )
SELECT subject_id,
  cohort_start_date,
  cohort_end_date,
  CAST (2 AS INT) AS cohort_definition_id
FROM #diagnoses;"
renderTranslateExecuteSql(conn, sql, cdm = "cdm_precision_test")

sql <- "TRUNCATE TABLE #diagnoses;
DROP TABLE #diagnoses;"
renderTranslateExecuteSql(conn, sql)
```

#### Covariate settings

```{r covariate settings, eval=TRUE}
covSettings <- createCovariateSettings(useDemographicsGender = TRUE,
                                    useDemographicsAge = TRUE,
                                    useDemographicsRace = TRUE,
                                    useDemographicsEthnicity = TRUE,
                                    useDemographicsIndexYear = TRUE,
                                    useConditionOccurrenceAnyTimePrior = TRUE,
                                    useConditionOccurrencePrimaryInpatientAnyTimePrior = TRUE,
                                    useConditionEraAnyTimePrior = TRUE,
                                    useConditionGroupEraAnyTimePrior = TRUE,
                                    useDrugExposureAnyTimePrior = TRUE,
                                    useDrugEraAnyTimePrior = TRUE,
                                    useDrugGroupEraAnyTimePrior = TRUE,
                                    useProcedureOccurrenceAnyTimePrior = TRUE,
                                    useDeviceExposureAnyTimePrior = TRUE,
                                    useMeasurementAnyTimePrior = TRUE,
                                    useMeasurementValueAnyTimePrior = TRUE,
                                    useMeasurementRangeGroupAnyTimePrior = TRUE,
                                    useObservationAnyTimePrior = TRUE,
                                    useCharlsonIndex = TRUE,
                                    useDcsi = TRUE,
                                    useChads2 = TRUE,
                                    useChads2Vasc = TRUE,
                                    useHfrs = TRUE,
                                    useDistinctConditionCountLongTerm = TRUE,
                                    useDistinctIngredientCountLongTerm = TRUE,
                                    useDistinctProcedureCountLongTerm = TRUE,
                                    useDistinctMeasurementCountLongTerm = TRUE,
                                    useDistinctObservationCountLongTerm = TRUE,
                                    useVisitCountLongTerm = TRUE,
                                    useVisitConceptCountLongTerm = TRUE,
                                    longTermStartDays = -365,
                                    endDays = 0)
```

#### Extract cohort data from cdm_precision_test schema

```{r extract cohort data, eval=TRUE}
plpData <- getPlpData(connectionDetails = cd,
                      cdmDatabaseSchema = "cdm_precision_test",
                      cohortDatabaseSchema = "cdm_precision_test",
                      cohortTable = "cohort",
                      cohortId = 1,
                      covariateSettings = covSettings,
                      outcomeDatabaseSchema = "cdm_precision_test",
                      outcomeTable = "cohort",
                      outcomeIds = 2)

summary(plpData)
```

#### Create a study population

```{r create study population, eval=TRUE}
population <- createStudyPopulation(plpData = plpData,
                                    outcomeId = 2,
                                    includeAllOutcomes = TRUE,
                                    firstExposureOnly = TRUE,
                                    washoutPeriod = 0,
                                    removeSubjectsWithPriorOutcome = FALSE,
                                    priorOutcomeLookback = 9999,
                                    requireTimeAtRisk = TRUE,
                                    minTimeAtRisk = 0,
                                    riskWindowStart = 0,
                                    riskWindowEnd = 365)
```

#### Making Predictions

```{r Predictions, eval=TRUE}
Prediction_COVID <- predictPlp(plpModel_COVID$model, population, plpData, index = NULL)
Prediction_COVIDDeath <- predictPlp(plpModel_COVID$model, population, plpData, index = NULL)
Prediction_COVIDVent <- predictPlp(plpModel_COVID$model, population, plpData, index = NULL)
```

#### Mapping Predictions to person_source_value

```{r Mapping Predictions, eval=TRUE}

#FileName is the csv file downloaded from PostgreSQL which has person_source_id mapped with person_id
FileName <- Person_table.csv 

CSV <- read.csv("FileName")

Result_COVID <- merge(Prediction_COVID,CSV,by.x="subjectId",by.y="person_id")
Result_COVIDDeath <- merge(Prediction_COVIDDeath,CSV,by.x="subjectId",by.y="person_id")
Result_COVIDVent <- merge(Prediction_COVIDVent,CSV,by.x="subjectId",by.y="person_id")

Result_COVID <- select(Result_COVID,"person_source_value","value")
Result_COVIDDeath <- select(Result_COVIDDeath,"person_source_value","value")
Result_COVIDVent <- select(Result_COVIDVent,"person_source_value","value")

write.csv(Final_result, "Submission_RandomForest_AllPatients_COVIDVent_AllPatients_AllPatients.csv")
```
